---
title: "cfChromatin analysis Sadeh et al. based"
author: "Mattia Zaghi"
date: "2024-04-04"
output: html_document
---


```{r setup 1}
.libPaths(c("/home/mattia/miniconda3/envs/cfChromatin/lib/R/library/"))
```

## Preprocessing of cfChromatin files ##

First upload the functions files necessary to perform the preprocessing steps 
```{r}
source("/home/mattia/cfChromatin_in_MS/cfChromatin_fun.R")
suppressMessages(library(rtracklayer))
suppressMessages(library(tools))
suppressMessages(library(ggplot2))
suppressMessages(library(parallel))
```
First we need to establish all the necessary directory to be used to process the files that we want to process

```{r}
catn = function(...) { cat(...,"\n") }
  setwd("/date/gcb/gcb_MZ/Analysis/cfChromatin_MS/")
  Mod = "H3K4me3"
  SourceDIR = "/date/gcb/gcb_MZ/Analysis/cfChIP-seq/"
  SetupDIR = paste0(SourceDIR, "SetupFiles/", Mod,"/")
  RootDir =  getwd()
  DataDir = paste0(RootDir, "/Samples/", Mod, "/")
  BedDir = paste0(RootDir, "/BED/", Mod, "/")
  TracksDIR = paste0(RootDir, "/Tracks/", Mod, "/")
  OutputDir = paste0(RootDir, "/Output/", Mod, "/")

  # Read the TSS windows data from an RDS file located in the SetupDIR directory.
  TSS.windows <- readRDS(file.path(SetupDIR, "Windows.rds"))

  # Filter to keep only standard chromosomes
  TSS.windows <- keepStandardChromosomes(TSS.windows)

  # Now TSS.windows contains only the TSS windows on standard chromosomes


  # Now TSS.windows contains only the TSS windows on standard chromosomes

  # Retrieve sequence information from the TSS windows data.
  genome.seqinfo = seqinfo(TSS.windows)
  # Define a list of chromosome names typically found in human genome datasets.
  ChrList = paste0("chr", c(1:22,"X","Y"))
  Win2Gene.Matrix.filename = paste0(SetupDIR,"Win2Gene.rds")
  print("Loading Window to Gene mapping")
  L = readRDS(Win2Gene.Matrix.filename)
  Win2Gene.matrix = L$Matrix
  MultiPromoterGenes = L$Multi
  rm(L)
#  colnames(Win2Gene.matrix)[is.na(colnames(Win2Gene.matrix))] = ""
  Genes = colnames(Win2Gene.matrix)
  GeneWindows = as.integer(rownames(Win2Gene.matrix))
  
  W = width(TSS.windows[GeneWindows])/1000
  GeneLength = ComputeGeneCounts(W)
  names(GeneLength) = Genes
  rm(W)
  
  Win2Gene.triplet = as(Win2Gene.matrix, "TsparseMatrix")
  Win2Gene.rows = Win2Gene.triplet@i +1
  Win2Gene.cols = Win2Gene.triplet@j +1
  Win2Gene.lists = split(Win2Gene.rows, Win2Gene.cols)
  names(Win2Gene.lists) = Genes
  rm(Win2Gene.cols)
  rm(Win2Gene.rows)
  rm(Win2Gene.triplet)
```

1. Find and upload the cfChromatin file coverage data from BED file to be subsequently analized

```{r}
# List all BED files in the directory
bed_files <- list.files(BedDir, pattern = "\\.bed$", full.names = TRUE)

# Define the parameters for cfChIP analysis
params <- cfChIP.Params()
params$DataDir <- BedDir  # Set the data directory
# ... set other parameters as needed ...

# Determine the number of cores to use
num_cores <- detectCores() - 1  # Leave one core free for other tasks

# Use parLapply to process BED files in parallel
# Iterate over the list of BED files and process each one
for(bed_file in bed_files) {
  # Print the current file being processed
  message("Processing file: ", bed_file)
  
  # Process the BED file and store the result
  processed_data <- cfChIP.ProcessFile(filename ="/date/gcb/gcb_MZ/Analysis/BED/H3K4me3/H001.1.tagAlign.gz", param = params)
  
  # Add the processed data to the list
  processed_data_list[[bed_file]] <- processed_data
}
# If you want to use a specific number of cores, replace num_cores with your desired number
# Now, 'processed_data_list' contains the processed data for each BED file
```
2. Normalize coverage from bed files using TSS catalogues 
```{r}

```

