---
title: "CfChromatin analysis script"
author: "Mattia Zaghi"
date: "2024-02-01"
output: html_document
---

```{r setup 1}
.libPaths(c("/home/mattia/R/x86_64-redhat-linux-gnu-library/4.3/"))
setwd(dir = "/home/mattia/cfChromatin_in_MS/")
```

```{r}
library(dplyr)
library(tidyverse)
library(ggpubr)
library(clipr)
library(GenomicRanges)
library(regioneR)
library(rtracklayer)
library(IRanges)
library(dplyr)
library(plyranges)
library(org.Hs.eg.db)
library(DESeq2)
library(Rsubread)
library(biomaRt)
library(AnnotationHub)
library(BSgenome.Hsapiens.UCSC.hg38)
library(parallel)
library(GenomeInfoDb)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(limma)
```

## 1.Constructing a TSS/Enhancers catalogue for cfChip-seq analysis 

As explained in the manual Page 5-6. The first step is creating a catalogue of regions with bonafide signal based on the ChromHMM annotation derived from 111 tissues abalyzed in the roadmap epigenomics project selecting only the genome portions annotated as TSS or TSS flanking regions.

```{r}
# Get the list of .bed files containing the specific ChromHMM annotations
files <- list.files("/date/gcb/gcb_MZ/Chrom_HMM_hg38_annotation/", pattern = "\\.bed", full.names = T)

# Load the .bed files into a list of Grange objects
encode <- lapply(files, read_bed)

# Rename the data frames in the encode list based on the file names
names(encode) <- sapply(files, function(file) {
  # Extract the base name of the file (without path)
  base_name <- basename(file)
  
  # Extract the first part of the name (e.g., "E001")
  new_name <- strsplit(base_name, "_")[[1]][1]
  
  # Return the new name
  return(new_name)
})

# Initialize an empty list to store the filtered data frames
encode_df <- list()

# Load the Glossary file which contains the code name corresponding to each Roadmap tissue
Glossary <- read_delim("/date/gcb/gcb_MZ/Chrom_HMM_hg38_annotation/Glossary.txt", 
delim = "\t", escape_double = FALSE, 
  trim_ws = TRUE)

# Convert the Glossary to a data frame
Glossary <-as.data.frame(Glossary)

# Loop over the encode list
for (i in 1:length(encode[1:111])){
  # Convert the list element to a data frame
  Tissue <- as.data.frame(encode[[i]]) 
  
  # Filter rows where 'name' is either "1_TssA" or "2_TssAFlnk"
  Tissue_TSS <- Tissue[(Tissue$name %in% c("1_TssA", "2_TssAFlnk")), ]
  
  # Convert the tissue in a Grange and reduce the range within each tissue
  Tissue_TSS<-makeGRangesFromDataFrame(Tissue_TSS,keep.extra.columns = T) %>% reduce()
  
  # Convert the GRanges object back to a data frame and store it in the list
  encode_df[[i]] <- as.data.frame(Tissue_TSS)

  # Get the current data frame
  df <- encode_df[[i]]
  
  # Get the EID from the name of the data frame
  eid <- names(encode)[i]
  
  # Get the corresponding STD_NAME from the Glossary data frame
  std_name <- Glossary$STD_NAME[Glossary$EID == eid]

  # Add the new column to the data frame
  df$tissue <- std_name
  
  # Store the modified data frame back in the list with the original name
  names(encode_df)[i] <- names(encode)[i]
  encode_df[[i]] <- df
}

# Combine all the merged regions from each tissue into a single data frame
ChromHMM_TSS <- do.call(rbind, encode_df)

# Group by seqnames, start, and end, and concatenate tissue values within each group
df <- df %>%
  group_by(seqnames, start, end) %>%
  summarise(tissue = paste(unique(tissue), collapse = ";"))

# Convert the data frame back to a GRanges object for further analysis
ChromHMM_TSS <- makeGRangesFromDataFrame(ChromHMM_TSS, keep.extra.columns = T) 
ChromHMM_TSS <- sort.GenomicRanges(ChromHMM_TSS)

# Merge overlapping ranges and concatenate the tissue names
gr <- reduce(ChromHMM_TSS , ignore.strand=TRUE)

gr<-join_overlap_inner(gr,ChromHMM_TSS)

# Group by seqnames, start, and end, and concatenate tissue values within each group
df <- as.data.frame(gr) %>%
  group_by(seqnames, start, end) %>%
  summarise(tissue = paste(unique(tissue), collapse = ";"))

# Convert the data frame back to a GRanges object for further analysis
Window <- makeGRangesFromDataFrame(df , keep.extra.columns = T) 

# Remove temporary variables to free up memory
rm(df)
rm(ChromHMM_TSS)
rm (gr)
rm(encode)
rm(encode_df)

```

The next step is to associate a gene to the as many as possible of those regions using the strategy that is explained in the manual which uses annotation from UCSC known genes and Ensembl to map to each one of those regions at list one transcript. 

```{r}

# Load the UCSC TSS list from the specified path
ucsc_genes<- read_tsv("/date/gcb/gcb_MZ/knownGene.txt", col_names = F) %>% 
  # Rename the columns for better readability
  dplyr::rename(Transcript_id=1,chr=2,strand=3,start=4,end=5) %>%
  # Select only the necessary columns
  dplyr::select(chr,start,end,strand,Transcript_id) 

# Define the list of chromosomes
chromosomes <- c("chr1", "chr2", "chr3", "chr4", "chr5", "chr6", "chr7", "chr8", "chr9", "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22", "chrX", "chrY", "chrM")

# Filter the UCSC genes to include only the defined chromosomes
ucsc_genes <- ucsc_genes[ucsc_genes$chr %in% chromosomes, ]

# Convert the data frame to a GRanges object
ucsc_genes <- makeGRangesFromDataFrame(ucsc_genes, keep.extra.columns = TRUE)



# Connect to the Ensembl database
ensembl <- useMart("ensembl")

# Select the human dataset from the Ensembl database
ensembl <- useDataset("hsapiens_gene_ensembl", mart = ensembl)

# Retrieve the transcript annotations from the Ensembl database
ensembl_transcripts <- getBM(attributes = c('ensembl_transcript_id', 'chromosome_name', 'start_position', 'end_position', 'strand', 'ensembl_gene_id', 'external_gene_name'), mart = ensembl)%>%
  # Replace "MT" with "M" in the chromosome names
  mutate(chromosome_name = ifelse(chromosome_name == "MT", "M", chromosome_name),
         # Add "chr" prefix to the chromosome names
         chromosome_name = paste0("chr", chromosome_name)) %>% 
   # Convert the strand values to "+" and "-"
   mutate(strand = ifelse(strand == 1, "+", 
                      ifelse(strand == -1, "-", strand))) %>% 
  # Rename the columns for better readability
  dplyr::rename(chr=2,start=3,end=4) 

# Filter the Ensembl transcripts to include only the defined chromosomes
ensembl_transcripts <- ensembl_transcripts[ensembl_transcripts$chr %in% chromosomes, ]

# Convert the data frame to a GRanges object
ensembl_transcripts <- makeGRangesFromDataFrame(ensembl_transcripts,keep.extra.columns = TRUE)

# Display the first few rows of the UCSC genes
head(ucsc_genes)
# Display the first few rows of the Ensembl transcripts
head(ensembl_transcripts)

```

The first step in order to associate a gene to each one of those regions is to find all the known TSS which are within 2.5 kb from the list of regions we have created using the collection from UCSC and Ensembl

```{r}
# Define the dimension for resizing the regions
distance <- 1


# Resize the gene or transcript regions from UCSC and Ensembl to the defined distance
# The resizing is done from the start of the regions
ucsc_genes_TSS<- resize(ucsc_genes, width = distance, fix="start")
ensembl_transcripts_TSS <-  resize(ensembl_transcripts, width = distance, fix="start")

# Find the overlaps between the TSS catalogue and the UCSC genes
# This will give us the annotations from UCSC that are nearby the TSS regions
Window_UCSC <- join_overlap_left(Window,ucsc_genes_TSS, maxgap = 2500) 

# After having annotated the TSS catalogue with UCSC we want to convert the transcript id into Gene names
transcript_ids <- mcols(Window_UCSC)$Transcript_id

# First, get the gene IDs corresponding to the transcript IDs
gene_ids <- mapIds(TxDb.Hsapiens.UCSC.hg38.knownGene, keys = transcript_ids, column = "GENEID", keytype = "TXNAME") 

# Get the number of cores
num_cores <- 40

# Create a function to process each element
process_element <- function(i) {
  # Get the current TranscriptID and GeneID(s)
  transcript_id <- names(gene_ids)[i]
  gene_id <- gene_ids[[i]]
  
  # If there are multiple GeneIDs, repeat the TranscriptID
  if (length(gene_id) > 1) {
    transcript_id <- rep(transcript_id, length(gene_id))
  }
  
  # Check if gene_id is empty
  if (length(gene_id) == 0) {
    # If gene_id is empty, assign it as NA
    gene_id <- NA
  }
  
  # Combine the current TranscriptID(s) and GeneID(s) into a data frame
  df_i <- data.frame(TranscriptID = transcript_id, GeneID = gene_id)
  
  return(df_i)
}

# Apply the function in parallel to each element in the list
df_list <- mclapply(seq_along(gene_ids), process_element, mc.cores = num_cores)

# Combine the list of data frames into one data frame
gene_ids_convert <- do.call(rbind, df_list)



# Then, map the gene IDs to gene symbols using org.Hs.eg.db

gene_symbols <- mapIds(org.Hs.eg.db, keys =gene_ids_convert$GeneID, column = "SYMBOL", keytype = "ENTREZID")

# Get the number of cores
num_cores <- 40

# Create a function to process each element
process_element <- function(i) {
  # Get the current TranscriptID and GeneID(s)
  Gene_id <- names(gene_symbols)[i]
  gene_symbols <- gene_symbols[[i]]
  
  # If there are multiple GeneIDs, repeat the TranscriptID
  if (length(gene_symbols) > 1) {
    transcript_id <- rep(Gene_id, length(gene_symbols))
  }
  
  # Check if gene_symbols is empty
  if (length(gene_symbols) == 0) {
    # If gene_symbols is empty, assign it as NA
    gene_symbols <- NA
  }
  
  # Combine the current TranscriptID(s) and GeneID(s) into a data frame
  df_i <- data.frame(GeneID = Gene_id, SYMBOL= gene_symbols)
  
  return(df_i)
}

# Apply the function in parallel to each element in the list
df_list <- mclapply(seq_along(gene_symbols), process_element, mc.cores = num_cores)

# Combine the list of data frames into one data frame
gene_symbols_convert <- do.call(rbind, df_list)

# Add the gene symbols to the metadata columns of the GRanges object
mcols(Window_UCSC)$name <- gene_symbols_convert$SYMBOL

#Now that we have the gene symbol associated with each UCSC annotation we can create a column name type in Which the annotated TSS will be names TSS_UCSC if annotated otherwise just TSS
Window_UCSC$type <- ifelse(!is.na(Window_UCSC$Transcript_id), "TSS_UCSC", "TSS")

                      
# Find the overlaps between the non-overlapping TSS regions and the Ensembl transcripts
# This will give us the annotations from Ensembl that are nearby the remaining TSS regions and not annotated using UCSC
Window_UCSC_ENSEMBL <- join_overlap_left(Window_UCSC,ensembl_transcripts_TSS, maxgap = 2500)

#We then convert the Grange into a dataframe to perform the final modification to the catalogue 
Window_UCSC_ENSEMBL <-as.data.frame(Window_UCSC_ENSEMBL)

#First in the column type all the TSS that have been annotated using ENSEMBL but not UCSC will be named TSS_ENSEMBL all the non annotated just TSS
Window_UCSC_ENSEMBL$type <- ifelse(!is.na(Window_UCSC_ENSEMBL$Transcript_id), "TSS_UCSC",
                  ifelse(!is.na(Window_UCSC_ENSEMBL$ensembl_transcript_id), "TSS_ENSEMBL", "TSS"))

# Use coalesce to fill NA values in 'name' with 'external_gene_name'
Window_UCSC_ENSEMBL$name <- coalesce(Window_UCSC_ENSEMBL$name, Window_UCSC_ENSEMBL$external_gene_name)

Window_UCSC_ENSEMBL[Window_UCSC_ENSEMBL == ""] <- NA

# Finally all the genomics windows with the same identical coordinates are merged together and the annotation are pull together in the different metadata column
Windows <- Window_UCSC_ENSEMBL %>%
  dplyr::select(-Transcript_id, -ensembl_transcript_id) %>%
  group_by(seqnames, start, end,type) %>%
  summarise(across(c(tissue, name), list(~paste(unique(.[!is.na(.)]), collapse = ";"))))
saveRDS(Windows, "Windows.rds")
```

To increase the number of annotated regions that are present in our catalouge now we will expand the region flanking our TSS to 3kb and try to find other overlapping TSS among the non assigned regions in the TSS catalouge

```{r}
# Increase the distance to 3kb for expanding the region flanking our TSS
distance<-3000

# Define TSS windows for each gene in UCSC and Ensembl by expanding the region flanking the TSS to 3kb
ucsc_tss_windows <- promoters(ucsc_genes_TSS, upstream = distance, downstream = distance)
ensembl_tss_windows <- promoters(ensembl_transcripts, upstream = distance, downstream = distance)

#To continue this operation we first need to retransfrom the windows dataset into a Grange object again
Windows<-readRDS("Windows.rds")
Windows<-makeGRangesFromDataFrame(Windows,keep.extra.columns = T)
# Find the overlaps between the non-annotated TSS regions and the UCSC TSS windows
# This will give us the non-annotated TSS regions that are nearby a UCSC gene
Windows_3Kb <- join_overlap_left(Windows,ucsc_tss_windows)

# Find the overlaps between the non-annotated TSS regions and the ENSEMBL TSS windows
# This will give us the non-annotated TSS regions that are nearby a ENSEMBL gene
Windows_3Kb <- join_overlap_left(Windows_3Kb,ensembl_tss_windows)

#Again transform the dataset into a df
Windows_3Kb<-as.data.frame(Windows_3Kb)
#Associate the correct TSS type to each line
Windows_3Kb$type <- ifelse(!is.na(Windows_3Kb$Transcript_id), "TSS_UCSC",
                  ifelse(!is.na(Windows_3Kb$ensembl_transcript_id), "TSS_ENSEMBL", "TSS"))

# Use coalesce to fill NA values in 'name' with 'external_gene_name'
Windows_3Kb$name <- coalesce(Windows_3Kb$name, Windows_3Kb$external_gene_name)

Windows_3Kb[Windows_3Kb == ""] <- NA

# Assuming 'Windows_3Kb' is your data frame
Windows_3Kb <-Windows_3Kb %>%
  dplyr::select(-Transcript_id, -ensembl_transcript_id) %>%
  group_by(seqnames, start, end,type) %>%
  summarise(across(c(tissue_1, name_1), list(~paste(unique(.[!is.na(.)]), collapse = ";"))))

Windows<-readRDS("Windows.rds")
# Use coalesce to fill NA values in 'name' with 'external_gene_name' and assocate the new tissue annotation to the old dataset
Windows$tissue_1<- coalesce(Windows$tissue_1, Windows_3Kb$tissue_1_1)
# Use coalesce to fill NA values in 'name' with 'external_gene_name' and associate the new name annotation to the old dataset
Windows$name_1<- coalesce(Windows$name_1, Windows_3Kb$name_1_1)

saveRDS(Windows,"Windows.rds")
```

The final step that now is missing is to add the the flanking and backgorund regions as described in the manual. Flakning regions of TSS will have a size of 1kb or lower. Background regions that are all the regions from the genome that not included in the TSS catalogue will be tiled into 5kb or lower windows and named as background for the subsequent analysis. 

```{r}
#first load the catalouge
Windows<-readRDS("Windows.rds")
#transform the Windows dataframe into a genomic ranges object again 

Windows<-makeGRangesFromDataFrame(Windows,keep.extra.columns = T)
# Create flanking regions at start and end of Windows dataset
flank_start <- IRanges::flank(Windows, 1000)
flank_end <-IRanges::flank(Windows,width = 1000,start = F)
#merge and sort them
flank<-c(flank_start,flank_end)
flank <- sort(flank)

# Name them as 'flanking'
mcols(flank)$type <- "flanking"
mcols(flank)$tissue_1 <- NA
mcols(flank)$name_1 <- NA
# Add the flanking regions to the Windows object
Windows<-c(Windows,flank)
Windows<-sort(Windows)


#The last step the needs to be performed is to tile the remaining regions at 5kb or lower and name them as background
# We first transform Windows into a Grange object named gr
gr <-Windows

# Get the sequence lengths for the genome of interest
seqinfo <- Seqinfo(genome="hg38")

# Create a GRanges object that represents the entire genome tiled at 5kb
genome_tiles <- tileGenome(seqinfo, tilewidth=5000, cut.last.tile.in.chrom=TRUE)

# Find the set difference between the genome tiles and your GRanges object
background <- setdiff_ranges(genome_tiles, gr)
background<-tile_ranges(background,width = 5000)
# Add the "background" type to these new regions
background$type <- "background"

# Combine your original GRanges object with the background regions and transform it in a dataframe
gr_with_background <- c(gr, background) 
gr_with_background <-sort.GenomicRanges(gr_with_background)%>% as.data.frame()

#finally the grange + background is reconverted into what will be the final Window object that we will use during our analysis
Windows<-gr_with_background %>% dplyr::select(seqnames,start,end,width,type,name_1,tissue_1) %>% dplyr::rename(name=6,tissue=7)

Windows<-makeGRangesFromDataFrame(Windows,keep.extra.columns = T)


#In the subsequent steps I realized that one range had a negative value so I use this code to modify it before going on

old_value <- '-799'  # replace this with the actual value

# Find the index of the range with the start value equal to 'old_value'
index <- which(start(Windows) == old_value)

# Change the start value of the range at 'index' to 1
start(Windows)[index] <- 1

saveRDS(Windows,"Windows.rds")

#N.B. the final windows object strucutre has been modelled after this dataset downloaded from the Sadeh et al. paper zenodo repo. you can check this dataset at this path "/date/gcb/gcb_MZ/Analysis/cfChIP-seq/SetupFiles/H3K4me3/Windows.csv" to see if I get the point
```


Now let´s create also a set of enhancers derived from ChromHMM annotation using “6_EnhG” and “7_Enh” marked regions to produce set of putative enhancers (N.B: this is still in development since I am using H3K27ac which should also map enhancers, but for now I am focusing my attention only on the widows TSS catalogue skip this part for now).

```{r}
# Initialize an empty list to store enhancer data
encode_enh<-list()

# Loop over each element in the 'encode' list
for (i in 1:length(encode)){
  # Convert the i-th element of 'encode' to a data frame
  Tissue<-as.data.frame(encode[[i]]) 

  # Filter rows where the 'name' column is either "6_EnhG" or "7_Enh"
  Tissue_enh<-Tissue[(Tissue$name %in% c("6_EnhG", "7_Enh")), ]

  # Add the filtered data frame to the 'encode_enh' list
  encode_enh[[i]] <- Tissue_enh

  # Set the name of the i-th element in 'encode_df' to the name of the i-th element in 'encode'
  names(encode_df)[i] <-  names(encode)[i]
}

# Combine all data frames in the 'encode_enh' list into a single data frame
ChromHMM_Ehn<- bind_rows(encode_enh)

# Convert the data frame to a GRanges object and reduce it to merge overlapping ranges
ChromHMM_Ehn<-makeGRangesFromDataFrame(ChromHMM_Ehn,keep.extra.columns = T) %>% reduce()

```

## 2.Constructing a Count Matrix of each chip-seq based on the Windows set of regions

 Now using the methodology described at page 7 of the manual we will construct a cont matrix of our chip samples over the windows catalouge.

```{r}

# Specify the directory where your BED files are stored
dir <- "/date/gcb/gcb_MZ/Round1/bed/bedtools/normal/"

# Get a list of all BED files in the directory that end with "H3K27ac_ChIP.bed"
bed_files <- list.files(path = dir, pattern = "H3K27ac_ChIP\\.bed$", full.names = TRUE)

#define regions
Windows<-readRDS("Windows.rds")

regions<- Windows

# Define a function to process a single BED file and a set of regions
process_bed_file <- function(bed_file, regions) {
  # Load the BED file
  bed <- import.bed(bed_file)

  # Resize the fragments to length 1, maintaining their center
  bed_resized <- resize(bed, width = 1, fix = "center")
  
  # Count overlaps
  counts <- countOverlaps(regions, bed_resized)

  # Add counts to the metadata columns of the regions GRanges object
  mcols(regions)$counts <- counts

  # Return the regions GRanges object
  return(regions)
}

# Use mclapply() to apply the function in parallel to the list of BED files
# Set mc.cores to the number of cores you want to use
results <- mclapply(bed_files, process_bed_file, regions = regions, mc.cores = 40)

# Now 'results' is a list of GRanges objects
# You can convert it to a named list of GRanges objects with the file names as the names of the list elements
counts_list<- setNames(results, gsub("\\.bed$", "", basename(bed_files)))

saveRDS(counts_list,"H3K27ac_ChIP_counts.rds")



```

## 3. Background Signal calculation

The background signal calculation is based on the premises and the methodology described at page 7-8-9 of the manual. Mine is simplified since I do not have cancer patients and also I am for now considering all the individuals as normal since I do not know what MS implies on this regards. 

```{r}
#load regions in which the background has been computed

counts_list<-readRDS("H3K27ac_ChIP_counts.rds")


#this function is based on the theoretical premises of the manual
estimateBackground <- function(gr) {
  # Extract background regions
  background_regions <- gr[gr$type == "background"]
  
  # Calculate the length of each region
  region_lengths <- width(background_regions)
  
  # Filter regions that are longer than 4kb
  long_background_regions <- background_regions[region_lengths >= 4000]
  
  # Extract counts from the filtered regions
  X <- mcols(long_background_regions)$counts
  
  # Find the 95th quantile of X
  T <- quantile(X, 0.95)
  
  # Restrict ourselves to values below T
  X <- X[X <= T]
  
  # Maximum likelihood of truncated poisson
  lambda_hat <- which.max(dpois(X, lambda = 1:length(X)))
  
  # Convert to reads/Kb (the median length of background windows is 5KB)
  lambda_hat <- lambda_hat / 5
  
  # Add the background estimate as a new column to the GRanges object
  gr$background <- lambda_hat
  
  return(gr)
}

# Apply the estimateBackground function to each GRanges object in the list
# Set mc.cores to the number of cores you want to use
granges_list_with_background <- mclapply(counts_list, estimateBackground, mc.cores = detectCores())

#Save the Grange object in which the bakcground computatio has been performed
saveRDS(granges_list_with_background,"H3K27ac_ChIP_counts.rds")

```

## 4. Gene-level signal and normalization to produce normalized tracks and counts for the rest of the analysis

This procedure follows what has been described in the manual at page 10- beginning of 11.

```{r}
#load the counts list needed to obtain the normalize signal in which the counts and the background signal has been already calculated

counts_list<-readRDS("H3K27ac_ChIP_counts.rds")

# Define the function to calculate raw signal at gene level
rawSignal <- function(gr) {
  # Extract counts and background from the GRanges object
  C <- mcols(gr)$counts
  B <- mcols(gr)$background * (width(gr)/1000)
  
  # Calculate raw signal based on what has been described in the manual. Therefore discard everything that is below the background level
  S <- ifelse(C > B, C - B, 0)
  
  # Add the raw signal as a new column to the GRanges object
  gr$rawSignal <- S
  
  return(gr)
}

# Apply the rawSignal function to each GRanges object in the list
rawSignal_counts <- mclapply(counts_list, rawSignal, mc.cores = detectCores())
rm(counts_list)

#In the list of my bed files there are these ChIP datasets from the literature. I am using them as QC to see how this script is impacting a conventional dataset, but I am not using this dataset for the scaling factor calculation since they are completely different from what I have in my dataset
exclude_samples <- c("CD19-GSE172116-1_H3K27ac_ChIP", "CD19-GSE172116-2_H3K27ac_ChIP")

# Remove the samples from the GRanges list
rawSignal_counts_blood_only <- rawSignal_counts[!names(rawSignal_counts) %in% exclude_samples]

# Extract the 'name' AKA genes and 'rawSignal' columns from the GRanges list
data <- do.call(rbind, lapply(rawSignal_counts_blood_only, function(gr) data.frame(name = gr$name, rawSignal = gr$rawSignal, stringsAsFactors = FALSE)))


# This operation is performed because later in the script I need to have the full list of the raw signal associated to each single region of my window catalouge
gr_list_modified <- lapply(rawSignal_counts_blood_only, function(gr) {
  gr$region <- paste0(seqnames(gr), ":", start(gr), "-", end(gr), "-", seq_along(gr))
  return(gr)
})

total_signal<-do.call(rbind, lapply(gr_list_modified, function(gr) data.frame(name = gr$region, rawSignal = gr$rawSignal, stringsAsFactors = FALSE)))

# Remove row names
rownames(total_signal) <- NULL


# Group by 'region' and summarize 'rawSignal'
total_signal<- total_signal %>%
  group_by(name) %>%
  summarize(rawSignal = sum(rawSignal, na.rm = TRUE))

total_signal<-as.data.frame(total_signal)


rownames(total_signal)<-total_signal$name


total_signal$name<-NULL

# Remove rows with NA in the 'name' column which correspond to the windows that have not been associated to any gene
data <- data[!is.na(data$name) & data$name != "", ] 

# Calculate the total signal for each gene across all samples
total_signal_gene <- tapply(data$rawSignal, data$name, sum)

# Calculate the standard deviation for each gene across all samples
std_dev <- tapply(data$rawSignal, data$name, sd)

# Calculate the median of the Poisson distribution (Since I did not know how to determine Housekeeping gene significance I set a casual value of the median on the poisson distribution and sd value across samples it can be modified to a strategy with more sense. I am open to any suggestion)
# Assuming 'lambda' is the mean of your data
lambda <- median(total_signal_gene)
poisson_median <- qpois(0.7, lambda)

# Define a standard deviation threshold
std_dev_threshold <- 5  # replace with your actual threshold

# Identify the housekeeping genes
# Housekeeping genes are those with a signal above the median of the Poisson distribution and a standard deviation below a certain threshold
housekeeping_genes <- names(total_signal_gene)[total_signal_gene > poisson_median & std_dev <= std_dev_threshold]

# Now we will first generated a matrix contained the counts for each specific housekeeping gene resulting from the sum of all the samples that then will be used for calculating the scaling factor
count_dfs <- list()

# Loop over the GRanges list
for(i in seq_along(rawSignal_counts_blood_only)) {
  # Extract the 'name' and 'counts' columns from the GRanges object
  data <- data.frame(name = rawSignal_counts_blood_only[[i]]$name, rawSignal = rawSignal_counts_blood_only[[i]]$rawSignal, stringsAsFactors = FALSE)
  
  # Group by 'name' and summarize 'counts'
  Windows <- data %>%
    group_by(name) %>%
    summarize(rawSignal = mean(rawSignal))
  
  # Remove rows with NA in the 'name' column
  All_genes <- Windows[!is.na(Windows$name) & Windows$name != "", ] 
  
  All_genes<-as.data.frame(All_genes)
  
  # Let's filter only the housekeeping genes that will be used in the analysis
  housekeeping <- All_genes %>% dplyr::filter(name %in% housekeeping_genes)
  
  # Set the 'name' column as row names
  row.names(housekeeping) <- housekeeping$name
  
  # Remove the 'name' column
  housekeeping$name <- NULL
  
  # Add the data frame to the list
  count_dfs[[i]] <- housekeeping
}
# Combine all data frames into a single count matrix
count_matrix <- do.call(cbind, count_dfs)

# Set the column names of the count matrix as the names of the GRanges objects
colnames(count_matrix) <- names(rawSignal_counts_blood_only)

#filter housekeeping genes from the gene count matrix

S_housekeeping <-as.matrix(count_matrix)

normalizeSignal <- function(gr_list, housekeeping_genes, reference_set) {

# I realize that in this part of the normalization there are issues that I would like to discuss with you directly to know what you think because I think this procedure is not fully correct but I do not know how to modify it at he moment 
# Loop over the GRanges list
for(i in seq_along(gr_list)) {
  # Apply quantile normalization to the raw signal of the housekeeping genes
  normalized_housekeeping_genes <- limma::normalizeQuantiles(S_housekeeping)
  
  # Extract the raw signal for all genes
  S_all <- as.matrix(total_signal)
  
  # Use linear regression to estimate a multiplicative normalization factor for each sample
  scaling_factor <- coef(lm(normalized_housekeeping_genes ~ S_housekeeping))[2]
  
  # Rescale the scaling factors so that the total normalized signal at the set of samples will be a million on average
  scaling_factor <- scaling_factor * 1e6 / scaling_factor * total_signal
  
 
  # Add the normalized signal as a metadata column of each Grange object by multiplying the raw signal for scaling factor 
  mcols(gr_list)$normalizedSignal <-  mcols(gr_list)$rawSignal * scaling_factor# replace 'column_name' with your actual column name
  
    }
    
    return(gr_list)
}


# replace with your actual housekeeping genes
reference_set <- names(rawSignal_counts_blood_only)  # replace with your actual reference set

# Apply the normalizeSignal function to each GRanges object in the list
normalizedSignal_counts <-normalizeSignal(gr_list = rawSignal_counts,housekeeping_genes = housekeeping_genes,reference_set = reference_set)


# Assuming 'gr_list' is your list of GRanges objects, 'old_name' is the old column name
# and 'new_name' is the new column name you want to use

# Apply the operation to all GRanges objects in the list
normalizedSignal_counts  <- lapply(gr_list, function(gr) {
  # Get the current column names
  col_names <- colnames(mcols(gr))
  
  # Find the index of the column you want to rename
  index <- which(col_names == "old_name")
  
  # Change the column name
  col_names[index] <- "new_name"
  
  # Assign the new column names back to the GRanges object
  colnames(mcols(gr)) <- col_names
  
  return(gr)
})


# Define a function to create a BigWig file from a GRanges object
createBigWig <- function(gr, filename) {
  # Create a GRanges object with the normalized signal as the score
  gr_with_score <- GRanges(seqnames = seqnames(gr),
                           ranges = ranges(gr),
                           strand = strand(gr),
                           score = mcols(gr)$normalizedSignal)
  
  # Export the GRanges object to a BigWig file
  export.bw(gr_with_score, filename)
}

# Apply the createBigWig function to each GRanges object in the list
filenames <- paste0("/date/gcb/gcb_MZ/Round1/R_bigiwig/", seq_along(normalizedSignal_counts), ".bw")  # replace with your actual filenames
mapply(createBigWig, normalizedSignal_counts, filenames)
```

