---
title: "CfChromatin analysis script"
author: "Mattia Zaghi"
date: "2024-02-01"
output: html_document
---

```{r setup 1}
.libPaths(c("/home/mattia/R/x86_64-redhat-linux-gnu-library/4.3/"))
```

```{r}
library(plyranges)
library(readr)
library(stringr)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(EnsDb.Hsapiens.v75)
library(org.Hs.eg.db)
library(parallel)
library(rtracklayer)
library(preprocessCore)
library(BSgenome.Hsapiens.UCSC.hg19)

sessionInfo()
```
#############################################################################################################################
###                                                          IMPORT                                                       ###
#############################################################################################################################
```{r}
source(paste0("/home/mattia/cfChromatin_in_MS/cfChromatin_fun.R"))
#working_path <- "/media/sf_ubuntu_music/cfChromatin/"
working_path <- "/date/gcb/gcb_MZ/"
annotation_dir_path <- paste0(working_path, "Chrom_HMM_hg19/")
glossary_path <- paste0(annotation_dir_path, "Glossary.txt")
bed_dir_path <- "/date/gcb/gcb_MZ/Round1/bed/bedtools/normal/"

core <- 6
wd<-"/date/gcb/gcb_MZ/Analysis/cfChIP-seq/SetupFiles/H3K27ac-MZ"
setwd(wd)

```
## 1.Constructing a TSS/Enhancers catalogue for cfChip-seq analysis 

As explained in the manual Page 5-6. The first step is creating a catalogue of regions with bonafide signal based on the ChromHMM annotation derived from 111 tissues abalyzed in the roadmap epigenomics project selecting only the genome portions annotated as TSS or TSS flanking regions.
#############################################################################################################################
###                                                    CATALOGUES CREATION                                                ###
#############################################################################################################################
```{r}
###### TSS
# Catalog part 1
windows <- load_catalog(annotation_dir = annotation_dir_path,glossary = glossary_path, glossary_group = "ANATOMY", states = c("1_TssA", "2_TssAFlnk"))
# Catalog part 2 & 3
windows <- annotate_catalog(windows, feature_tag = "TSS", db = list('UCSC'=TxDb.Hsapiens.UCSC.hg19.knownGene, 'Ensembl'=EnsDb.Hsapiens.v75), db_gap = 2500, db_spe = TRUE, db_tag = "EXTRA_TSS", db_spe_center = 3000)
# Catalog part 4
windows <- add_flank(windows, feature_tag = "FLANKING", flanking_size = 1000)
windows <- add_background(windows, feature_tag = "BACKGROUND", bin_size = 5000)

# Statistics
show_statistics(windows, db_tag = "EXTRA_TSS")

# Save files
saveRDS(windows,paste0(wd,"Windows.rds"))

df_windows <- data.frame(
  seqnames=seqnames(windows),
  starts=start(windows)-1,
  ends=end(windows),
  tissue=windows$tissue,
  feature=windows$feature,
  type=windows$type)

write.table(df_windows, file=paste0(wd,"Windows.bed"), quote=F, sep=",", row.names=F, col.names=F)
```
## 2.Constructing a Count Matrix of each chip-seq based on the Windows set of regions

 Now using the methodology described at page 7 of the manual we will construct a cont matrix of our chip samples over the windows catalouge.

#############################################################################################################################
###                                                 PROCESSING SEQUENCING FILES                                           ###
#############################################################################################################################
```{r}
### Process fastq files separetly to get BED files

histone_mark <- "H3K27ac"

# Get a list of all BED files in the directory that end with "H3K27ac_ChIP.bed"
bed_files <- list.files(path = bed_dir_path, pattern = paste0(histone_mark,"_ChIP\\.bed$"), full.names = TRUE)
# Define regions
windows <- readRDS(paste0(working_path,"TSS_windows.rds"))
# Use mclapply() to apply the function in parallel to the list of BED files
# Set mc.cores to the number of cores you want to use
results <- mclapply(bed_files, process_bed_file, regions = windows, mc.cores = core)
# Now 'results' is a list of GRanges objects
# You can convert it to a named list of GRanges objects with the file names as the names of the list elements
counts_list <- setNames(results, gsub("\\.bed$", "", basename(bed_files)))

saveRDS(counts_list, paste0(working_path,paste0(histone_mark,"_counts.rds")))
```
## 3. Background Signal calculation

The background signal calculation is based on the premises and the methodology described at page 7-8-9 of the manual. Mine is simplified since I do not have cancer patients and also I am for now considering all the individuals as normal since I do not know what MS implies on this regards. 
#############################################################################################################################
###                                                 ESTIMATING BACKGROUNG SIGNAL                                          ###
#############################################################################################################################
```{r}
#counts_list <- readRDS(paste0(working_path,paste0(histone_mark,"_counts.rds")))

# Apply the estimateBackground function to each GRanges object in the list
# Set mc.cores to the number of cores you want to use
granges_list_with_background <- mclapply(counts_list, estimateBackground, tag = "BACKGROUND", background_cutOff = 4000, quantile_cutOff = 0.95, mc.cores = core)

saveRDS(granges_list_with_background, paste0(working_path,paste0(histone_mark,"_counts_bg.rds")))
```
## 4. Gene-level signal and normalization to produce normalized tracks and counts for the rest of the analysis

This procedure follows what has been described in the manual at page 10- beginning of 11.

#############################################################################################################################
###                                                     FEATURE LEVEL SIGNAL                                              ###
#############################################################################################################################
```{r}
# Remove the samples from the GRanges list

#granges_list_with_background_rm <- granges_list_with_background[!names(granges_list_with_background) %in% exclude_samples]

# Apply the calculationSignal function to each GRanges object in the list
# Set mc.cores to the number of cores you want to use
granges_list_geneSignal <- mclapply(granges_list_with_background, calculationSignal, tag_list = c("TSS","EXTRA_TSS"), rm_feature_list=c("UNKNOWN"), mc.cores = core)
df_geneSignal <- t(as.data.frame(do.call(rbind, granges_list_geneSignal)))

saveRDS(granges_list_geneSignal, paste0(working_path,paste0(histone_mark,"_signal.rds")))
saveRDS(df_geneSignal, paste0(working_path,paste0(histone_mark,"_df_signal.rds")))
```
##############################################################################################################
###                                                    CALCULATE SIZE FACTOR                                           ###
#############################################################################################################################
```{r}
###############
#df_geneSignal <- as.data.frame(readRDS(paste0(histone_mark,"_df_signal.rds")))

# Create different dataframe for each type of sample
plasma_sample <- c("12097-P_H3K27ac_ChIP", "14223-P_H3K27ac_ChIP", "16057-P-100_H3K27ac_ChIP", "16057-P-200_H3K27ac_ChIP", "16170-P_H3K27ac_ChIP", "19019-P_H3K27ac_ChIP")
csf_sample <- c("04061-C-100_H3K27ac_ChIP", "04061-C-200_H3K27ac_ChIP", "12097-C_H3K27ac_ChIP", "14223-C_H3K27ac_ChIP", "16170-C_H3K27ac_ChIP", "19019-C_H3K27ac_ChIP")
df_plasma <- df_geneSignal[,plasma_sample]
df_csf <- df_geneSignal[,csf_sample]

# Calculate the size factor
sf_plasma <- calculation_sf(df_plasma, ctrl=c("19019-P_H3K27ac_ChIP"), housekeeping=1000)
sf_csf <- calculation_sf(df_csf, ctrl=c("19019-C_H3K27ac_ChIP"), housekeeping=1000)

# Apply size factor to genes
df_plasma_geneSignal <- t(t(df_geneSignal[,names(sf_plasma)])*sf_plasma)
df_csf_geneSignal <- t(t(df_geneSignal[,names(sf_csf)])*sf_csf)

# Apply size factor to windows
sf_joined <- c(sf_plasma,sf_csf)
for(x in names(sf_joined)){
  no_bg <- granges_list_with_background[[x]]$counts-granges_list_with_background[[x]]$background
  # Remove background signal from each window and apply the size factor
  granges_list_with_background[[x]]$norm <- ifelse(no_bg > 0,no_bg*sf_joined[x],0)
}

saveRDS(df_plasma_geneSignal, paste0(working_path,paste0(histone_mark,"_plasma_df_signal_norm.rds")))
saveRDS(df_csf_geneSignal, paste0(working_path,paste0(histone_mark,"_csf_df_signal_norm.rds")))
saveRDS(granges_list_with_background, paste0(working_path,paste0(histone_mark,"_counts_norm.rds")))

```
