---
title: "Human_Single_cell_H3K27ac_integration"
author: "Mattia Zaghi"
date: "2024-05-29"
output: html_document
---

```{r setup 2}
library(Signac)
library(Seurat)
library(GenomicRanges)
library(future)
library(stringr)
library(ggplot2)
library(gghalves)
library(ggpubr)
library(dplyr)
library(tidyverse)
library(EnsDb.Hsapiens.v86)
library(BSgenome.Hsapiens.UCSC.hg38)
library(ComplexUpset)
library(regioneR)
library(scales)
library(ggvenn)
library(readxl)
library(ggthemes)
library(dplyr)
library(tidyverse)
library(plyranges)
library(pbapply)
```


First We need to load all the datasets that we need to use for the bridge integration


```{r}
Human_pairTag_Cortex<-readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined.obj_Human_Droplet_PairTag_all_mod.rds")

Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@meta.data$barcode<-paste(Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@meta.data$sample,Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@meta.data$barcode_RNA,sep = "_")
Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@meta.data$barcode <- sub("H3K27ac_", "", Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@meta.data$barcode)



Human_pairTag_Cortex_multiome<- CreateSeuratObject(
  counts = Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@assays[["RNA"]]$counts,
  assay = "RNA"
)
Human_pairTag_Cortex_multiome[["SCT"]]<-Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@assays[["SCT"]]
Human_pairTag_Cortex_multiome[["RNA"]]<- Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@assays[["RNA"]]
Human_pairTag_Cortex_multiome[["H3K27ac_bins"]]<-Human_pairTag_Cortex[["H3K27ac_CCTATCCT"]]@assays[["H3K27ac_bins"]]

metadata_RNA<-as.data.frame(Human_pairTag_Cortex[["RNA_AAAAGGGG_H3K27ac"]]@meta.data) %>% unique()
metadata_RNA$barcode<-paste(metadata_RNA$sample,metadata_RNA$barcode_RNA,sep = "_")
# Assuming df is your data frame and 'column_name' is the name of your column
metadata_RNA$barcode <- sub("H3K27ac_", "", metadata_RNA$barcode)

Human_pairTag_Cortex_multiome@meta.data$barcode<-colnames(Human_pairTag_Cortex_multiome)
barcode<-as.data.frame(Human_pairTag_Cortex_multiome@meta.data$barcode) %>% dplyr::rename(barcode=1)
metadata_RNA<-inner_join(barcode,metadata_RNA)
metadata_RNA<-metadata_RNA[1:36]
metadata_RNA<-unique(metadata_RNA) %>% dplyr::rename(predicted.id=15)
rownames(metadata_RNA)<-metadata_RNA$barcode
Human_pairTag_Cortex_multiome@meta.data<-metadata_RNA

metadata_H3K27ac<-as.data.frame(Human_pairTag_Cortex[["H3K27ac_CCTATCCT"]]@meta.data)
metadata_H3K27ac<-metadata_H3K27ac[1:43]
metadata_H3K27ac$barcode<-paste(metadata_H3K27ac$sample,metadata_H3K27ac$barcode_RNA,sep = "_")
metadata_H3K27ac<-inner_join(barcode,metadata_H3K27ac) %>% unique()
metadata<-inner_join(metadata_H3K27ac,metadata_RNA,by="barcode")
rownames(metadata)<-metadata$barcode

Human_pairTag_Cortex_multiome@meta.data<-metadata

barcode<-as.character(Human_pairTag_Cortex_multiome@meta.data[["barcode"]])
colnames(Human_pairTag_Cortex_multiome)<-barcode
Human_pairTag_Cortex_multiome_n<-RenameCells(
Human_pairTag_Cortex_multiome,
new.names = barcode)

DefaultAssay(Human_pairTag_Cortex_multiome) <- "RNA"
Human_pairTag_Cortex_multiome <- NormalizeData(Human_pairTag_Cortex_multiome, normalization.method = "LogNormalize", scale.factor = 10000)
Human_pairTag_Cortex_multiome<- FindVariableFeatures(Human_pairTag_Cortex_multiome, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(
Human_pairTag_Cortex_multiome)

Human_pairTag_Cortex_multiome  <- ScaleData(
Human_pairTag_Cortex_multiome , features = all.genes)

# store mitochondrial percentage in object meta data
Human_pairTag_Cortex_multiome <- PercentageFeatureSet(Human_pairTag_Cortex_multiome , pattern = "^MT-", col.name = "percent.mt")

DefaultAssay(Human_pairTag_Cortex_multiome) <- "SCT"

Human_pairTag_Cortex_multiome<- RunPCA(Human_pairTag_Cortex_multiome, features = VariableFeatures(object = Human_pairTag_Cortex_multiome),reduction.name = "pca_SCT")

Human_pairTag_Cortex_multiome<- FindNeighbors(Human_pairTag_Cortex_multiome, dims = 1:15,reduction = "pca_SCT")
Human_pairTag_Cortex_multiome <- FindClusters(Human_pairTag_Cortex_multiome, resolution = 0.5,
                                                                  cluster.name="Seurat_clusters_SCT")
Human_pairTag_Cortex_multiome <- RunUMAP(Human_pairTag_Cortex_multiome, dims = 1:15, reduction = "pca_SCT",reduction.name = "umap_SCT")

DefaultAssay(Human_pairTag_Cortex_multiome) <- "RNA"

Human_pairTag_Cortex_multiome <- RunPCA(Human_pairTag_Cortex_multiome , features = VariableFeatures(object = Human_pairTag_Cortex_multiome))
Human_pairTag_Cortex_multiome <- FindNeighbors(Human_pairTag_Cortex_multiome, dims = 1:15)
Human_pairTag_Cortex_multiome<- FindClusters(Human_pairTag_Cortex_multiome, resolution = 0.5)

Human_pairTag_Cortex_multiome<- RunUMAP(Human_pairTag_Cortex_multiome, dims = 1:15)

pc <- 10

#Human_pairTag_Cortex_multiome<-readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/Human_pairTag_Cortex_multiome_object.rds")
DefaultAssay(Human_pairTag_Cortex_multiome) <- "H3K27ac_bins"
Human_pairTag_Cortex_multiome <- FindTopFeatures(Human_pairTag_Cortex_multiome, min.cutoff = 10)
Human_pairTag_Cortex_multiome <- RunTFIDF(Human_pairTag_Cortex_multiome)
Human_pairTag_Cortex_multiome<- RunSVD(Human_pairTag_Cortex_multiome)
Human_pairTag_Cortex_multiome <- RunUMAP(Human_pairTag_Cortex_multiome, reduction= 'lsi', dims = 2:pc)
Human_pairTag_Cortex_multiome<- FindNeighbors(Human_pairTag_Cortex_multiome,reduction = 'lsi', dims = 2:pc)
Human_pairTag_Cortex_multiome<- FindClusters(Human_pairTag_Cortex_multiome, verbose = FALSE, algorithm = 3,resolution = 0.5)


# Get the column names that start with "prediction.score"
cols <- grep("^prediction.score", colnames(Human_pairTag_Cortex_multiome@meta.data), value = TRUE)

# Remove these columns from the metadata
Human_pairTag_Cortex_multiome@meta.data <- Human_pairTag_Cortex_multiome@meta.data[, !colnames(Human_pairTag_Cortex_multiome@meta.data) %in% cols]


```
Before proceeding we need to map our Human pair Tag dataset onto our qeury dataset

```{r}
# load developing embryonic brain reference atlas
Human_CTX_atlas<- readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/ref/Siletti_CX_EA_seurat_ENS.rds")

# Assuming 'seurat_object' is your Seurat object
cell_counts <- table(Human_CTX_atlas@meta.data[["supercluster_term"]])

# Print the cell counts
print(cell_counts)


# Assuming 'seurat_object' is your Seurat object and 'cell_population' is your metadata column
set.seed(123) # for reproducibility

# Get the unique cell populations
cell_populations <- unique(Human_CTX_atlas@meta.data[["supercluster_term"]])

# Initialize an empty vector to store the downsampled cell names
downsampled_cells <- c()

# Loop over each cell population
for (pop in cell_populations) {
  # Get the cells in the current population
  cells_in_pop <- which(Human_CTX_atlas@meta.data[["supercluster_term"]] == pop)
  
  # If the population has more than 1000 cells, downsample it
  if (length(cells_in_pop) > 1000) {
    cells_in_pop <- sample(cells_in_pop, size = 1000)
  }
  
  # Add the cells to the downsampled cells vector
  downsampled_cells <- c(downsampled_cells, cells_in_pop)
}

# Subset the Seurat object to only include the downsampled cells
Human_CTX_atlas_downsampled <- subset(Human_CTX_atlas, cells = downsampled_cells)

# Assuming 'seurat_object' is your Seurat object
cell_counts <- table(Human_CTX_atlas_downsampled@meta.data[["supercluster_term"]])

# Print the cell counts
print(cell_counts)
rm(Human_CTX_atlas)

gc()

# transfer cell type labels from reference to query
transfer_anchors <- FindTransferAnchors(reference =Human_CTX_atlas_downsampled , query = Human_pairTag_Cortex_multiome,
                                            reference.assay = "RNA",
                                        reduction = "cca",
                                        npcs = 50,
                                        dims = 1:50,
  query.assay = "RNA",
  features = VariableFeatures(Human_pairTag_Cortex_multiome@assays[["RNA"]]))
  

supercluster <- TransferData(
  anchorset = transfer_anchors, 
  refdata =  Human_CTX_atlas_downsampled@meta.data[["supercluster_term"]],
  weight.reduction = Human_pairTag_Cortex_multiome@reductions[["pca"]],
  dims = 1:50
)

cell_type <- TransferData(
  anchorset = transfer_anchors, 
  refdata =  Human_CTX_atlas_downsampled@meta.data[["cell_type"]],
  weight.reduction = Human_pairTag_Cortex_multiome@reductions[["pca"]],
  dims = 1:50,
)

supercluster_obj<-Human_pairTag_Cortex_multiome
cell_type_obj<-Human_pairTag_Cortex_multiome

supercluster_obj<- AddMetaData(
  object = Human_pairTag_Cortex_multiome,
  metadata = supercluster
)

cell_type_obj <- AddMetaData(
  object = Human_pairTag_Cortex_multiome,
  metadata = cell_type
)


# If the dimensions are correct, assign them to the metadata

  Human_pairTag_Cortex_multiome@meta.data[["supercluster"]] <- supercluster_obj@meta.data[["predicted.id"]]
  
  # Get the column names that start with "prediction.score"
  cols <- grep("^prediction.score", colnames(supercluster_obj@meta.data), value = TRUE)
  
  # Loop over these column names
for (col in cols) {
  # Create new column name by replacing "prediction.score" with "supercluster.score"
  new_col <- gsub("prediction.score", "supercluster.score", col)
  
  # Add new column to Human_pairTag_Cortex_multiome@meta.data
  Human_pairTag_Cortex_multiome@meta.data[[new_col]] <- supercluster_obj@meta.data[[col]]
}

  Human_pairTag_Cortex_multiome@meta.data[["cell_type"]] <- cell_type_obj@meta.data[["predicted.id"]]
  
    # Get the column names that start with "prediction.score"
  cols <- grep("^prediction.score", colnames(cell_type_obj@meta.data), value = TRUE)
  
  # Loop over these column names
for (col in cols) {
  # Create new column name by replacing "prediction.score" with "supercluster.score"
  new_col <- gsub("prediction.score", "cell_type.score", col)
  
  # Add new column to Human_pairTag_Cortex_multiome@meta.data
  Human_pairTag_Cortex_multiome@meta.data[[new_col]] <- cell_type_obj@meta.data[[col]]
}
  
saveRDS(Human_pairTag_Cortex_multiome,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/Human_pairTag_Cortex_multiome_object.rds")
```

Map scH3K27ac dataset using bridge integration
Now that we have the reference, query, and bridge datasets set up, we can begin integration. The bridge dataset enables translation between the scRNA-seq reference and the scATAC-seq query, effectively augmenting the reference so that it can map a new data type. We call this an extended reference, and first set it up. Note that you can save the results of this function and map multiple scATAC-seq datasets without having to rerun.

First, we drop the first dimension of the ATAC reduction

```{r}
dims.atac <- 2:50
dims.rna <- 1:50



Human_pairTag_Cortex_multiome<-readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/Human_pairTag_Cortex_multiome_object.rds")


Human_CTX_atlas<- readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/ref/Siletti_CX_EA_seurat_ENS.rds")

# Assuming 'seurat_object' is your Seurat object
cell_counts <- table(Human_CTX_atlas@meta.data[["supercluster_term"]])

# Print the cell counts
print(cell_counts)


# Assuming 'seurat_object' is your Seurat object and 'cell_population' is your metadata column
set.seed(123) # for reproducibility

# Get the unique cell populations
cell_populations <- unique(Human_CTX_atlas@meta.data[["supercluster_term"]])

# Initialize an empty vector to store the downsampled cell names
downsampled_cells <- c()

# Loop over each cell population
for (pop in cell_populations) {
  # Get the cells in the current population
  cells_in_pop <- which(Human_CTX_atlas@meta.data[["supercluster_term"]] == pop)
  
  # If the population has more than 1000 cells, downsample it
  if (length(cells_in_pop) > 1000) {
    cells_in_pop <- sample(cells_in_pop, size = 1000)
  }
  
  # Add the cells to the downsampled cells vector
  downsampled_cells <- c(downsampled_cells, cells_in_pop)
}

# Subset the Seurat object to only include the downsampled cells
Human_CTX_atlas_downsampled <- subset(Human_CTX_atlas, cells = downsampled_cells)

# Assuming 'seurat_object' is your Seurat object
cell_counts <- table(Human_CTX_atlas_downsampled@meta.data[["supercluster_term"]])

# Print the cell counts
print(cell_counts)
rm(Human_CTX_atlas)

gc()

DefaultAssay(Human_pairTag_Cortex_multiome) <-  "SCT"
DefaultAssay(Human_CTX_atlas_downsampled) <- "SCT"
obj.rna.ext <- PrepareBridgeReference(
  reference = Human_CTX_atlas_downsampled , bridge = Human_pairTag_Cortex_multiome,
  reference.reduction = "pca", reference.dims = dims.rna,bridge.query.assay = "H3K27ac_bins",
  normalization.method = "SCT")
#saveRDS(obj.rna.ext,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/bridge_integration_human_CTX_H3K27ac.rds")
rm(Human_CTX_atlas_downsampled)

gc()
```
Now, we can directly find anchors between the extended reference and query objects. We use the FindBridgeTransferAnchors function, which translates the query dataset using the same dictionary as was used to translate the reference, and then identifies anchors in this space. The function is meant to mimic our FindTransferAnchors function, but to identify correspondences across modalities.


```{r}
rm(Human_pairTag_Cortex_multiome_n)

gc()
pc <- 10

Human_CZI_Cortex_SC<-readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined.obj_Human_CZI_nanoCT.rds")
#Human_CZI_Cortex_SC<-Human_CZI_Cortex_SC[["H3K27ac_TATAGCCT"]]
#Human_CZI_Cortex_SC <- RunUMAP(Human_CZI_Cortex_SC,reduction = 'lsi', dims = 2:pc)
#Human_CZI_Cortex_SC<- FindNeighbors(Human_CZI_Cortex_SC,reduction = 'lsi', dims = 2:pc)
#Human_CZI_Cortex_SC<- FindClusters(Human_CZI_Cortex_SC, verbose = FALSE, algorithm = 3,resolution = 0.5)

bridge.anchor <- FindBridgeTransferAnchors(
  extended.reference = obj.rna.ext, query = Human_CZI_Cortex_SC,
  reduction = "lsiproject", dims = dims.atac, bridge.reduction = "cca")


```
Once we have identified anchors, we can map the query dataset onto the reference. The MapQuery function is the same as we have previously introduced for reference mapping . It transfers cell annotations from the reference dataset, and also visualizes the query dataset on a previously computed UMAP embedding. Since our reference dataset contains cell type annotations at three levels of resolution (l1 - l3), we can transfer each level to the query dataset.


```{r}
# Map the 'supercluster_term' metadata
supercluster_mapped <- MapQuery(
  anchorset = bridge.anchor, 
  reference = obj.rna.ext,
  query =  Human_CZI_Cortex_SC,
  refdata = "supercluster_term")

# Map the 'cell_type' metadata
cell_type_mapped <- MapQuery(
  anchorset = bridge.anchor, 
  reference = obj.rna.ext,
  query =  Human_CZI_Cortex_SC,
  refdata = "cell_type")

  Human_CZI_Cortex_SC@meta.data[["supercluster"]] <- supercluster_mapped@meta.data[["predicted.id"]]

  Human_CZI_Cortex_SC@meta.data[["supercluster.score.max"]] <- supercluster_mapped@meta.data[["predicted.id.score"]]

saveRDS(Human_CZI_Cortex_SC,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined.obj_Human_CZI_nanoCT.rds")

rm(supercluster_mapped)
rm(cell_type_mapped)

gc()
```

Now we can visualize the results, plotting the scATAC-seq cells based on their predicted annotations, on the reference UMAP embedding. You can see that each scATAC-seq cell has been assigned a cell name based on the scRNA-seq defined cell ontology.


```{r fig.width = 10, fig.height = 10}
p1<-DimPlot(
  Human_CZI_Cortex_SC, group.by = "supercluster",
  reduction = "umap", label = TRUE
) + ggtitle("H3K27ac") +NoLegend()

p2<-DimPlot(
  Human_CZI_Cortex_SC, group.by = "cell_type",
  reduction = "umap", label = TRUE
) + ggtitle("Predictions") +NoLegend()

p1+p2
```

Now let's try to do data integration using gene activity and then label transfer in the classical using siletti Et al. cortex dataset


```{r}
# Assuming `multiomic_data` is your multiomic Seurat object and `h3k27ac_data` is your H3K27ac-only Seurat object
DefaultAssay(Human_pairTag_Cortex_multiome)<-"H3K27ac_bins"
DefaultAssay(Human_CZI_Cortex_SC)<-"H3K27ac_bins"
# Step 1: Data Integration
anchors <- FindIntegrationAnchors(object.list = list(Human_pairTag_Cortex_multiome,Human_CZI_Cortex_SC), dims = 1:30)
integrated_data <- IntegrateData(anchorset = anchors, dims = 1:30)

# Step 2: Cell Similarity
integrated_data <- ScaleData(integrated_data)
integrated_data <- RunPCA(integrated_data)
integrated_data <- FindNeighbors(integrated_data, dims = 1:30)
integrated_data <- FindClusters(integrated_data, resolution = 0.5)

# Step 3: Infer Gene Expression
# This is a simplified example. In practice, you might want to use a more sophisticated method.
for (cluster in unique(Idents(integrated_data))) {
  cells_in_cluster <- WhichCells(integrated_data, idents = cluster)
  average_expression <- AverageExpression(Human_pairTag_Cortex_multiome[["RNA"]], cells = cells_in_cluster)
  # Add the average expression to the Human_CZI_Cortex_SC dataset
  Human_CZI_Cortex_SC[["RNA"]][, WhichCells(Human_CZI_Cortex_SC, idents = cluster)] <- average_expression
}


saveRDS(Human_CZI_Cortex_SC,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined.obj_Human_CZI_nanoCT.rds")

```


Now let´s try do a data integration between Paired-Tag and the CZI nanoCT

```{r}

Human_CZI_Cortex_SC@meta.data[["barcode"]]<-paste0(Human_CZI_Cortex_SC@meta.data[["sample"]],"_",Human_CZI_Cortex_SC@meta.data[["barcode"]])
# first add dataset-identifying metadata
Human_pairTag_Cortex_multiome$dataset <- "PairTag"
Human_CZI_Cortex_SC$dataset <- "nanoCT"


# merge
Brain.combined <- merge(Human_pairTag_Cortex_multiome, Human_CZI_Cortex_SC)

# process the combined dataset
DefaultAssay(Brain.combined) <- "H3K27ac_bins"
Brain.combined <- FindTopFeatures(Brain.combined, min.cutoff = 10)
Brain.combined <- RunTFIDF(Brain.combined)
Brain.combined <- RunSVD(Brain.combined)

saveRDS(Brain.combined,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined_human_CNS_H3K27ac.rds")

```

Now let´s perform the data integration between the two single cell H3K27ac

try to armonize better the two datasets


```{r}
Brain.combined<-readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined_human_CNS_H3K27ac.rds")

Brain.combined<-RunUMAP(Brain.combined, reduction = "lsi", dims = 2:30)
#unintegrated UMAP looking not that good
DimPlot(Brain.combined, reduction = "umap", group.by = c("dataset"))
```

let's try to integrate together the 2 datasets


```{r}
Brain.combined <- IntegrateLayers(object = Brain.combined, method = CCAIntegration, orig.reduction = "lsi", new.reduction = "integrated.cca",
    verbose = FALSE)

# re-join layers after integration
ifnb[["RNA"]] <- JoinLayers(ifnb[["RNA"]])

ifnb <- FindNeighbors(ifnb, reduction = "integrated.cca", dims = 1:30)
ifnb <- FindClusters(ifnb, resolution = 1)
```



```{r}
Brain.combined<-readRDS("/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/combined_human_CNS_H3K27ac.rds")
# find integration anchors
DefaultAssay(Human_pairTag_Cortex_multiome) <- "H3K27ac_bins"
DefaultAssay(Human_CZI_Cortex_SC) <- "H3K27ac_bins"
integration.anchors <- FindIntegrationAnchors(
  object.list = list(Human_pairTag_Cortex_multiome,Human_CZI_Cortex_SC),
  anchor.features = rownames(Human_pairTag_Cortex_multiome),
  reduction = "rlsi",
  dims = 2:30
)

# integrate LSI embeddings
integrated <- IntegrateEmbeddings(
  anchorset = integration.anchors,
  reductions =Brain.combined@reductions[["lsi"]],
  new.reduction.name = "integrated_lsi",
  dims.to.integrate = 1:30
)

# Identify anchors between the RNA-seq and ATAC-seq datasets
anchors <- FindTransferAnchors(reference = Human_pairTag_Cortex_multiome, query = Human_CZI_Cortex_SC, features = VariableFeatures(Human_pairTag_Cortex_multiome@assays[["H3K27ac_bins"]]))

# Transfer the data
predictions <- TransferData(anchorset = anchors, refdata = multiome_seurat_obj@assays$RNA@counts, weight.reduction = atac_seurat_obj[["H3K27ac"]], dims = 2:30)

# create a new UMAP using the integrated embeddings
integrated <- RunUMAP(integrated, reduction = "integrated_lsi", dims = 2:30)
saveRDS(integrated,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/integrated_human_CNS_H3K27ac.rds")

integrated@meta.data[["sample.x"]]<-NULL
integrated@meta.data[["sample.x"]]<-NULL

sample<- data.frame(sample=colnames(integrated))

rownames(sample)<-sample$sample

sample$sample<- sub("_[A-Z]*-[0-9]*$", "", sample$sample)
sample$barcode<- rownames(sample)
pairtagbarcodes <- sample$barcode[grepl("^M1", sample$barcode)]
nanoCTbarcodes <- sample$barcode[grepl("NanoCT", sample$barcode)]

integrated@meta.data[["sample.x"]]<-NULL
integrated@meta.data[["sample.x"]]<-NULL
integrated@meta.data[["sample"]]<-sample$sample

integrated  <- subset(integrated, subset = cell_type != "central nervous system macrophage")
integrated  <- subset(integrated, subset = cell_type != "endothelial cell")
integrated<- subset(integrated, subset = cell_type != "fibroblast")
integrated<- subset(integrated, subset = cell_type != "pericyte")

integrated_subset_pairtag = subset(integrated, cells=pairtagbarcodes)

integrated_subset_nanoCT = subset(integrated, cells=nanoCTbarcodes)

p1<-DimPlot(integrated_subset_pairtag,group.by="supercluster")
p2<-DimPlot(integrated,group.by="cell_type")

integrated_subset_supercluster = subset(integrated, subset = supercluster.score.max  > 0.7)
# Assuming 'seurat_obj' is your Seurat object, 'column_name' is the name of the column,
# and 'entry' is the specific entry you want to exclude
integrated_subset_supercluster  <- subset(integrated_subset_supercluster, subset = cell_type != "central nervous system macrophage")
integrated_subset_supercluster  <- subset(integrated_subset_supercluster, subset = cell_type != "endothelial cell")
integrated_subset_supercluster  <- subset(integrated_subset_supercluster, subset = cell_type != "fibroblast")

integrated_subset_celltype = subset(integrated, subset = cell_type.score.max > 0.9)
saveRDS(integrated_subset,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/integrated_human_CNS_H3K27ac_subset.rds")
p1 <- DimPlot(integrated_subset_supercluster, group.by = "cell_type")
p2 <- DimPlot(integrated_subset_celltype, group.by = "cell_type")
p1
p2
```
Let´s plot some marker genes to check that the cellular Identity that have been obtained are the expected ones. 



```{r,fig.height=10, fig.width=20}
DefaultAssay(integrated) <- "RNA"
FeaturePlot(integrated, features = c("OLIG2", "SOX10", "SOX9","RBFOX3","MBP","MAG","PDGFRA","OPALIN","GFAP","ALDH1L1",
                  "PLP1","AQP4","TUBB3","MAP2","TMEM119","CX3CR1"))
```

Let´s perform a dotplot to evaluate the expression of the "marker genes" inside each transferred cell population.


```{r}
#now I will check all the markers by subetting x dataset



DefaultAssay(integrated_subset_supercluster) <- "RNA"
# Assuming `seurat_object` is your Seurat object and `gene_markers` is a vector of gene markers
gene_markers <- c("OLIG2", "SOX10", "SOX9","RBFOX3","MBP","MAG","PDGFRA","OPALIN","GFAP","ALDH1L1",
                  "PLP1","AQP4","TUBB3","MAP2","TMEM119","CX3CR1") # replace with your actual gene markers

# Create a dot plot
DotPlot(integrated, features = gene_markers,group.by = "cell_type") + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5))

# Create a violin plot
VlnPlot(integrated_subset_supercluster, features = "CXCR3", group.by = "supercluster")

```


The final step is call peaks in the single cell dataset, resize them to 400bp and then try to understand which one of them is enrich in which cell type. Basically a peakset relative to each cell type our annotated object. 


```{r}
library(Signac)
library(Seurat)
library(GenomicRanges)

# Load your Seurat object
# pbmc <- readRDS("path_to_your_data.rds")
DefaultAssay(integrated_subset)<-"H3K27ac_bins"
# Call peaks
peaks <- CallPeaks(
  object =integrated_subset,
  group.by = "cell_type",macs2.path = "/home/mattia/miniconda3/envs/macs2/bin/macs2"
)

# Resize the peaks
resized_peaks <- resize(peaks, width = 400, fix = "center")

# Assuming 'peaks' is your GRanges object
filtered_peaks <- subset(resized_peaks, seqnames(resized_peaks) %in% c(paste0("chr", 1:22), "chrX", "chrY"))

# Drop unused levels
seqlevels(filtered_peaks) <- droplevels(seqlevels(filtered_peaks))


write_bed(peaks,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/integrated_human_CNS_H3K27ac.peaks.bed")

# Create a binary peak-by-cell matrix
peak_matrix <- FeatureMatrix(
  fragments = integrated_subset@assays[["H3K27ac_bins"]]@fragments,
  features = filtered_peaks,
  cells = colnames(integrated_subset)
)
annotations <- GetGRangesFromEnsDb(ensdb = EnsDb.Hsapiens.v86)

# Assuming 'peak_matrix' is your matrix with the called peaks
chrom_assay <- CreateChromatinAssay(counts = peak_matrix,sep = c(":", "-"),annotation = annotations)


# Add the ChromatinAssay to your existing Seurat object
integrated_subset[["peaks"]] <-chrom_assay

saveRDS(integrated_subset,"/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/integrated_human_CNS_H3K27ac_subset.rds")

DefaultAssay(integrated_subset) <- "peaks"

Idents(integrated_subset) <- "cell_type"

marker_peaks <- FindAllMarkers(
  object = integrated_subset,
  only.pos = TRUE,
  min.pct = 0.1,
  logfc.threshold = 0.25,
  group.by = "cell_type"
)


```

Extract BigWig files from clusters to check if the peaks calls make sense 



```{r}

ExportGroupBW(
  object = integrated_subset, #The seurat object
  assay =  "H3K27ac_bins", #The assay of the fragment file
  group.by = "cell_type", #The metadata used to split the fragment file
  idents = NULL, #The idents from the group.by to be exported
  normMethod = "RC",#Normalization method for the biwig files. It can be the name of any quantitative column in the @meta.data or ncells as the number of cells. Can also be set to none
  tileSize = 100, #The size of the tiles in the bigwig file
  minCells = 5, #The minimum of cells in a group to be exported
  cutoff = 4, #The maximum number of fragment in a given tile
  outdir =  "/date/gcb/gcb_MZ/Analysis/Single_Cell_data_mining/Tracks", #The output directory for bigwig files. Also used as output directory for SlitFragment function
  verbose=TRUE #Display messages
)


```