---
title: "Create TSS window for cfChIP"
author: "Mattia Zaghi"
date: "2024-02-01"
output: html_document
---

```{r setup 1}
.libPaths(c("/home/mattia/R/x86_64-redhat-linux-gnu-library/4.3/"))
```

```{r}
library(dplyr)
library(tidyverse)
library(ggpubr)
library(clipr)
library(GenomicRanges)
library(regioneR)
library(rtracklayer)
library(IRanges)
library(dplyr)
library(plyranges)
library(org.Hs.eg.db)
library(DESeq2)
library(Rsubread)
library(biomaRt)
library(AnnotationHub)
```

## Constructing a TSS/Enhancers catalogue for cfChip-seq analysis 

First we need to load the whole catalogue fo encode tissue. We will use the regions that have been lifted over to Hg38 genome.

```{r}
files <- list.files("/date/gcb/gcb_MZ/Chrom_HMM_hg38_annotation/", pattern = "\\.bed", full.names = T)
encode <-  lapply(files, read_bed)
```

To create a cataolog of regions that might act as ground truth for H3K27ac we select all the regions that were marked with states "1_TssA", "2_TssAFlnk”, “6_EnhG” and “7_Enh” to include both TSS and enhancers which are the regions that are mostly considered positive for this type of marker. 

First let´s start by selecting and filtering active TSS

```{r}
encode_df<-list()

for (i in 1:length(encode)){
  Tissue<-as.data.frame(encode[[i]]) 
  Tissue<-Tissue[(Tissue$name %in% c("1_TssA", "2_TssAFlnk")), ]

    encode_df[[i]] <- Tissue
    names(encode_df)[i] <-  names(encode)[i]
}

TSS<- bind_rows(encode_df)

TSS<-makeGRangesFromDataFrame(TSS,keep.extra.columns = T) %>% reduce()
```

The we Need to interpolate this list of TSS with UCSC list and Ensembl to validate the ones that have been annotated 

```{r}

# Path to UCSC TSS list
TSS_UCSC <- read_tsv("/date/gcb/gcb_MZ/knownGene.txt", col_names = F) %>% 
  dplyr::rename(Transcript_id=1,chr=2,strand=3,start=4,end=5) %>%
  dplyr::select(chr,start,end,strand,Transcript_id) %>% 
  makeGRangesFromDataFrame(keep.extra.columns = T)

# Display the first few rows of the BED file
head(TSS_UCSC)

```
```{r}
ah = AnnotationHub()
dm <- query(ah, c( "ENSEMBL", "Homo Sapiens"))
f<-dm[["AH5046"]]
```

```{r}

listMarts()

# Connect to the ENSEMBL Biomart

ensembl <- useMart("ENSEMBL_MART_ENSEMBL", host = 'www.ensembl.org', path = '/biomart/martservice', dataset = 'hsapiens_gene_ensembl')


# Get TSS positions for ENSEMBL transcripts in hg38
ENSEMBL_TSS <- getBM(attributes = c("chromosome_name", "start_position", "end_position", "strand", "ensembl_transcript_id"), mart = ensembl) %>% 
  dplyr::rename(chr=2,start=3,end=4) 
  ENSEMBL_TSS$chr<-paste0("chr", ENSEMBL_TSS$chr) 
  ENSEMBL_TSS$strand<- ifelse(ENSEMBL_TSS$strand == 1, "+", "-")
  ENSEMBL_TSS$chr<- ifelse(ENSEMBL_TSS$chr== "chrMT", "chrM",ENSEMBL_TSS$chr)
ENSEMBL_TSS<-makeGRangesFromDataFrame(ENSEMBL_TSS,keep.extra.columns = T)

# Print the first few rows of the resulting data
head(ENSEMBL_TSS)
```
First we will look for overlap between our list of putative TSS and the UCSC known gene list and ENSEMBL transcipts list 
```{r}
TSS_overlap_UCSC <- find_overlaps_within(TSS,TSS_UCSC,maxgap = 2500)




TSS_ENSEMBL<-find_overlaps_within(TSS,ENSEMBL_TSS,maxgap = 2500)

TSS_ENSEMBL_unique <-subsetByOverlaps(TSS_ENSEMBL,TSS_overlap_UCSC,invert = T)
# Concatenate the two GRanges objects
TSS_overlap_UCSC <-as.data.frame(TSS_overlap_UCSC)
TSS_overlap_UCSC <-TSS_overlap_UCSC[1:12]
TSS_overlap_UCSC <-TSS_overlap_UCSC %>% dplyr::rename(itemRgb=8,name=6,score=7,ensembl_transcript_id=12)

TSS_first_selection <- rbind.data.frame(TSS_overlap_UCSC,TSS_non_overlapping)

head(TSS_first_selection)
```

